---
title: "Inspecting the Militarized Dispute Locations dataset"
author: "Ian Roberts"
date: "April 6, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(message = F)
knitr::opts_chunk$set(warning = F)
```

```{r mapYear, cache = TRUE}
library(leaflet)
library(Hmisc)
mdl <- read.csv("c:/Users/Ian/Downloads/midloc-v1-1-MIDLOC_1.1.csv", header = T)
mdl <- mdl[-grep(-9999.000000, mdl$latitude),]

mdl$cutyear <- cut2(x = mdl$year, g = 5)

groups <-  levels(mdl$cutyear)
groupColors <-  colorFactor(palette = rainbow(n = 5), domain = mdl$cutyear)

mapY = mdl %>% 
        leaflet() %>% 
          addTiles()
        
for(i in 1:5){
mapY = mapY %>% addCircleMarkers(lat = subset(mdl, cutyear == groups[i])$latitude, 
                         lng = subset(mdl, cutyear == groups[i])$longitude,
                         clusterOptions = markerClusterOptions(),
                         label = subset(mdl, cutyear == groups[i])$midlocmeasuringpoint,
                         color = groupColors(subset(mdl, cutyear == groups[i])$cutyear),
                         group = groups[i])
}

mapY %>% 
        addLayersControl(overlayGroups = groups) %>% 
        addLegend(position = "topright", 
                  pal = groupColors, 
                  values = mdl$cutyear)
```

```{r mapPrecision, cache=TRUE}
library(leaflet)
#mdl <- read.csv("c:/Users/Ian/Downloads/midloc-v1-1-MIDLOC_1.1.csv", header = T)
#mdl <- mdl[-grep(-9999.000000, mdl$latitude),]

groups <- levels(as.factor(mdl$precision))
groupColors <- colorFactor(palette = rainbow(n = 5), domain = as.factor(mdl$precision))

mapP = mdl %>% 
        leaflet() %>% 
        addTiles()

for(i in 1:5){
        mapP = mapP %>% addCircleMarkers(lat = subset(mdl, precision == groups[i])$latitude, 
                                         lng = subset(mdl, precision == groups[i])$longitude,
                                         clusterOptions = markerClusterOptions(),
                                         label = subset(mdl, precision == groups[i])$midlocmeasuringpoint,
                                         color = groupColors(subset(mdl, precision == groups[i])$precision),
                                         radius = subset(mdl, precision == groups[i])$precision^2,
                                         group = groups[i])
}

mapP %>% 
        addLayersControl(overlayGroups = groups) %>% 
        addLegend(position = "topright", 
                  pal = groupColors, 
                  values = c("1: Contested Point", 
                             "2: Approximate Contested Point", 
                             "3: Subnational Contested Area", 
                             "4: Contested Border/River/Lake", 
                             "5: Contested National Area"))